\documentclass[12pt]{report}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
%\usepackage{palatino}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{amsfonts}
\usepackage{calrsfs}

\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}[theorem]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]

\numberwithin{equation}{theorem}

\begin{document}
\input{title}

\newpage
\section*{Acknowledgement of Sources}

For all ideas taken from other sources (books, articles, internet),
the source of the ideas is mentioned in the main text and fully
referenced at the end of the report.
All material which is quoted essentially word-for-word from
other sources is given in quotation marks and referenced.
Pictures and diagrams copied from the internet or other sources
are labelled with a reference to the web page or book, article etc.

\textbf{Signed}: Jishnu Kaiwar

\textbf{Date}: \today


\tableofcontents

\chapter{Introduction}
\label{chap:intro}
\input{intro}

\chapter{The Nielsen Schreier Theorem}
\label{chap:ns}
\input{ns}

\chapter{The Word Problem}
\label{chap:wp}
\input{wp}

\chapter{Conclusions}
\label{chap:conc}

The main results that we have discussed in this project are the Nielsen-Schreier Theorem and the Novikov-Boone-Britton Theorem, which led us to the study of simplical complexes and Turing machines. The chief result regarding simplical complexes was \ref{thm:covering-fundamental} with which we find that subgroups of fundamental groups are fundamental groups of covering spaces. These results and all have more general topological statements gives us much insight into both how fundamental groups are structured as well as how topological spaces can be covered. Investigating further along this line leads us to the \emph{universal cover} for any topological space which is the covering space associated with the trivial subgroup $\{ 1 \}$ of a fundamental group with many interesting properties \cite{hatcher2002AT}.

The second part of the project dealt with the word problem for groups and semigroups. We introduced the notion of decision processes informally and briefly discussed how this can be defined into formal mathematics. To do this we introduced the Turing machine, a model for the universal computer. We proceeded to prove the Markov-Post theorem that dealt with the word problem for semigroups and finally talked about the Novikov-Boone-Britton theorem that we didn't prove fully, omitting the proof of Boone's Lemma. Topics of the discussion here led to the development of the field of computability theory and laid the foundations for theoretical computer science. A mathematical study of this field can be found in \cite{bridges1994}

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

%%% Local Variables:
%%% TeX-master: "master"
%%% End: